% LaTeX file for resume 
% This file uses the resume document class (res.cls)

\documentclass{res} 
%\usepackage{helvetica} % uses helvetica postscript font (download helvetica.sty)
%\usepackage{newcent}   % uses new century schoolbook postscript font 
\usepackage{amsmath,amssymb,amstext,amsgen,amsbsy,amsopn,amsfonts,graphicx,amsthm,cleveref}

%\usepackage{parskip}
\usepackage[left=3cm, right=3cm, top=2.5cm, bottom=4cm,footskip=1cm,headsep=1.5cm]{geometry}
%example of geometry package \geometry{verbose,tmargin=3cm,bmargin=3.5cm,lmargin=3.5cm,rmargin=3.5cm,headheight=3cm,headsep=4cm,footskip=2cm}
\usepackage[all]{xy}
\usepackage[OT2,T1]{fontenc}
%\usepackage[latin9]{inputenc}
\usepackage{titletoc}
\setlength{\textheight}{9.5in} % increase text height to fit on 1-page 

\begin{document} 

\name{\textbf{Curriculum Vitae: Adam Morgan}\\[25pt]}     % the \\[12pt] adds a blank
				        % line after name      

%\address{\bf  CONTACT INFORMATION \\Max Planck Institute for Mathematics\\
%Vivatsgasse 7\\
%53111 Bonn\\
%Germany }
\address{E-mail: ajmorgan44@gmail.com \\Date of birth: October 24, 1989\\ Webpage: amorgan516.github.io}
                                  
\begin{resume}       
 
 \section{EMPLOYMENT}       

\bigskip 

\vspace{-0.1in}	
   \begin{tabbing}
   \hspace{2in}\= \hspace{2in}\= \kill % set up two tab positions
    {\bf University of Glasgow}\>~\> ~~~~~~ {\bf Oct 2021- }\\
   \end{tabbing}\vspace{-20pt}      % suppress blank line after tabbing
     EPSRC Research Fellow

\vspace{-0.1in}	
   \begin{tabbing}
   \hspace{2in}\= \hspace{2in}\= \kill % set up two tab positions
    {\bf Mathematisches Forschungsinstitut Oberwolfach}\>~\> ~~~~~~ {\bf Jan 2021- April 2021 }\\
   \end{tabbing}\vspace{-20pt}      % suppress blank line after tabbing
     Leibniz  fellow  

\vspace{-0.1in}	
   \begin{tabbing}
   \hspace{2in}\= \hspace{2in}\= \kill % set up two tab positions
    {\bf MPIM Bonn}\>~\> ~~~~~~ {\bf Oct 2019- Oct 2020}\\
   \end{tabbing}\vspace{-20pt}      % suppress blank line after tabbing
      Postdoctoral fellow  

\vspace{-0.1in}	
   \begin{tabbing}
   \hspace{2in}\= \hspace{2in}\= \kill % set up two tab positions
    {\bf University of Glasgow}\>~\> ~~~~~~ {\bf Oct 2018 - Sept 2019}\\
   \end{tabbing}\vspace{-20pt}      % suppress blank line after tabbing
      Research Associate   

\vspace{-0.1in}	
   \begin{tabbing}
   \hspace{2in}\= \hspace{2in}\= \kill % set up two tab positions
    {\bf King's College London}\>~\> ~~~~~~ {\bf Sept 2016 - Sept 2018}\\
   \end{tabbing}\vspace{-20pt}      % suppress blank line after tabbing
      Research Associate   
   
\vspace{-0.1in}	
   \begin{tabbing}
   \hspace{2in}\= \hspace{2in}\= \kill % set up two tab positions
    {\bf University of Warwick}\>~\> ~~~~~~ {\bf Oct 2015 - Aug 2016}\\
   \end{tabbing}\vspace{-20pt}      % suppress blank line after tabbing
      Research Assistant    
      
\vspace{-0.1in}	
   \begin{tabbing}
   \hspace{2in}\= \hspace{2in}\= \kill % set up two tab positions
    {\bf Detica}\>~\> ~~~~~~ {\bf June - Sept 2011}\\
   \end{tabbing}\vspace{-20pt}      % suppress blank line after tabbing
      Intern (Programming and statistical modelling, primarily in SAS)    
      
\section{EDUCATION}          
        
\bigskip

\vspace{-0.1in}	
   \begin{tabbing}
   \hspace{2in}\= \hspace{2in}\= \kill % set up two tab positions
    {\bf University of Bristol}\\
 ~\\
      {\it PhD in Mathematics} \>~\> ~~~~~~ {\bf Oct 2012 - Sept 2015} \\       
    Supervised by Professor Tim Dokchitser      
  \end{tabbing}\vspace{-17pt}      % suppress blank line after tabbing
     \begin{tabbing}
   \hspace{2in}\= \hspace{2in}\= \kill % set up two tab positions
    {\bf Clare College, University of Cambridge}\\
~\\
        % suppress blank line after tabbing
      {\it MMath (Part III of the Mathematical Tripos) } \>~\> ~~~~~~ {\bf Oct 2011 - June 2012}\\       
    Passed with Distinction (ranked 5th) \\   
      % suppress blank line after tabbing
~\\
      {\it BA in Mathematics } \>~\> ~~~~~~ {\bf Oct 2008 - June 2011} \\       
    First class each year   \\   
\end{tabbing}\vspace{-25pt}   



\section{GRANTS AND PRIZES} 
\bigskip 
EPSRC postdoctoral fellowship, awarded 2021 \medskip \\
    Faculty of Science Commendation for excellence in a doctoral thesis awarded by University of Bristol, 2015 \medskip \\
    EPSRC doctoral training grant, 2012 - 2016 \medskip \\         
    Robert Greene cup for academic excellence awarded by Clare College, June 2012 \medskip \\
Horne Prize for Physical Sciences awarded by Clare College, three consecutive years 2009-2011 \medskip \\
Salters Horners advanced physics prizewinner 2008 (top mark in A-Level physics)      

\section{ACCEPTED PAPERS}         
        
\bigskip

\vspace{-0.1in}	
{\it A note on hyperelliptic curves with ordinary reduction over 2-adic fields}, with Vladimir Dokchitser, J. Number Theory 244 (2023), 264-278\

\vspace{-0.1in}	
{\it Arithmetic of hyperelliptic curves over local fields}, with Tim Dokchitser, Vladimir Dokchitser and C\'{e}line Maistret, Math. Ann. (2022), doi: 10.1007/s00208-021-02319-y  \

{\it A user's guide to the local arithmetic of hyperelliptic curves}, with A. Best, L.A. Betts, M. Bisatt, R. van Bommel, V. Dokchitser, O. Faraggi, S. Kunzweiler, C. Maistret, S. Muselli, S. Nowell,  Bull. Lond. Math. Soc. 54 (2022), no. 3, 825--867

{\it On 2-Selmer groups of twists after quadratic extension}, with Ross Paterson, J. London Math. Soc. 105 (2022), no. 2, 1110-1166. \

{\it Tate module and bad reduction}, with Tim Dokchitser and Vladimir Dokchitser, Proc. Amer. Math. Soc. 149 (2021), 1361-1372.

{\it Quadratic twists of abelian varieties and disparity in Selmer ranks},  Algebra Number Theory 13 (2019), no. 4, 839–899. \

{\it Semistable types of hyperelliptic curves}, with Tim Dokchitser, Vladimir Dokchitser and C\'{e}line Maistret, Algebraic curves and their applications, 73–135, Contemp. Math., 724, Amer. Math. Soc., Providence, RI, 2019. \

   {\it Integral module structure of $\Lambda_{A/K}$ for Jacobians of
semistable hyperelliptic curves of genus $2$}, with Vladimir Dokchitser, appendix to L. Alexander Betts and Vladimir Dokchitser, {\it Finite quotients of  $\mathbb{Z}[C_n]$-lattices and Tamagawa numbers of semistable abelian varieties}, Math. Proc. Cambridge Philos. Soc. 166 (2019), no. 3, 487–521. 
%
\section{PREPRINTS}         
        
\bigskip

{\it Parity of ranks of Jacobians of curves}, with V. Dokchitser, H. Green and A. Konstantinou, arXiv:2211.06357   

{\it Field change for the Cassels-Tate pairing and applications to class groups}, with Alexander Smith, arXiv:2206.13403 . \

{\it The Cassels--Tate pairing for finite Galois modules}, with Alexander Smith, arxiv:2103.08530 . \

{\it The 4-rank of class groups of $K(\sqrt{n}) $}, with Peter Koymans and Harry Smit, arxiv:2101.03407. \

{\it Isogenies between abelian varieties with good ordinary reduction}, 3 pages, appendix to  Vladimir Dokchitser and C\'{e}line Maistret, {Parity conjecture for abelian surfaces},  arxiv:1911.04626. \

{\it $2$-Selmer parity for hyperelliptic curves over quadratic extensions},  arXiv:1504.01960. (updated April 2022)\


\section{INVITED LECTURE COURSES}

\bigskip
\vspace{-0.1in}
Four lecture course {\it Local arithmetic of curves and Jacobians} for {\it CMI-HIMR Summer school in Computational Number Theory}, University of Bristol, June, 2019.

Four lecture course {\it L-functions and the Birch and Swinnerton-Dyer conjecture} for the summer school {\it Curves, L-functions and Galois representations}, ICTP Trieste, September 2017.\

 \section{ORGANISATIONAL ACTIVITIES}
 
 \medskip
 
 \textup{Organiser of \textit{Gauge Fields in Arithmetic, Topology and Physics}},   \textup{ICMS Edinburgh, March 6-10 2023}  
 
\textup{Organiser of \textit{Young Researchers in Algebraic Number Theory} (Y-RANT) 2022},   \textup{University of Glasgow, Aug 23-25}  
 
\textup{Organiser of the  Algebra and Number Theory Seminar, }   \textup{University of Glasgow, Oct 2021-}\medskip\\
\textup{Organiser for the number theory study group} {\it Average ranks of elliptic curves,} \textup{University of Warwick, Sept-Dec 2015.} 

\section{SELECTED TALKS}         
        
 
{\it  On the Hasse principle for Kummer varieties}, Number Theory Seminar, {\it The Hebrew University of Jerusalem}, June 2023    \medskip  \\
{\it  On the Hasse principle for Kummer varieties}, invited speaker at the 33rd meeting of London-Paris Number Theory Seminar, Imperial College London, June 2023    \medskip  \\
{\it  Counterexamples to the Hasse principle in quadratic twist families of genus 1
hyperelliptic curves}, invited speaker at {\it Number Theory Section, Representation Theory XVIII}, Dubrovnik, June 2023    \medskip  \\
{\it  Invariants of hyperelliptic curves over local fields}, invited speaker at {\it Number Theory meets Tropical Geometry: the semistable reduction of curves}, MPIM Leipzig, April 2023    \medskip  \\
{\it  Hasse principle for Kummer varieties in the case of generic 2-torsion}, invited speaker at {\it Arithmetic, Algebra, and Algorithms}, ICMS Edinburgh, April 2023    \medskip  \\
{\it  Parity of ranks of abelian varieties}, invited speaker for the Algebraic Number Theory and Diophantine Geometry workshop at the British Mathematical Colloquium 2023 \medskip  \\
{\it  Hasse principle for Kummer varieties in the case of generic 2-torsion}, Heilbronn seminar, University of Bristol, March 2023    \medskip  \\
{\it  Hasse principle for Kummer varieties in the case of generic 2-torsion}, Number Theory Seminar, University of Oxford, Feb 2023    \medskip  \\
{\it  Parity of ranks of abelian varieties}, invited speaker at {\it Mordell 2022: a centenary of elliptic curves}, University of Cambridge, Aug 2022    \medskip  \\
{\it  Integral Galois module structure of Mordell--Weil groups}, Queen Mary Number Theory Seminar, Dec 2021   \medskip  \\
{\it  Integral Galois module structure of Mordell-Weil groups}, invited speaker at the Paris-London Number Theory Seminar, Université Sorbonne Paris Nord, Nov 2021  \medskip \\
{\it  The $2$-parity conjecture for Jacobians of hyperelliptic curves}, Glasgow Algebra and Number Theory Seminar, Oct 2021    \medskip  \\
{\it  Integral Galois module structure of Mordell--Weil groups}, invited speaker at {\it Arithmetic Statistics and Local Global Principles}, ESI Vienna Sept 2021    \medskip  \\
{\it  Integral Galois module structure of Mordell--Weil groups}, invited speaker at {\it Workshop on arithmetic statistics problems}, June 2021   \medskip  \\
{\it  Invariants of hyperelliptic curves over local fields}, KTH Number Theory seminar, Stockholm, Feb 2020  \vspace{-0.18in} \medskip  \\
{\it Parity of Selmer ranks in quadratic twist families}, Intercity seminar, Leiden, Feb 2020 \medskip \\
{\it Parity of Selmer ranks in quadratic twist families}, Number theory seminar, MPIM Bonn, Dec 2019 \medskip \\
{\it Parity of Selmer ranks in quadratic twist families}, Warwick Number Theory seminar, University of Warwick, June 2019 \medskip \\
{\it Class groups, Selmer groups and Cassels--Tate pairings}, invited speaker at the conference {\it Enumerative Arithmetic and the Cohen–Lenstra Heuristics}, MPIM, Bonn, June 2019 \medskip \\
{\it Parity of Selmer ranks in quadratic twist families}, London Number Theory seminar, University College London, January 2019  \medskip \\
{\it Parity of ranks of abelian varieties}, Algebra seminar, University of Glasgow, November 2018  \medskip \\
{\it Parity of ranks of abelian varieties},  contributed talk for the conference {\it Young Researchers in Algebraic Number Theory}, University of Sheffield, November 2018  \medskip \\
{\it Parity of 2-Selmer ranks of abelian varieties over quadratic extensions},  contributed talk for the conference {\it Rational and Integral Points via Analytic and Geometric Methods}, CMO Oaxaca, June 2018  \medskip \\
{\it Parity of 2-Selmer ranks of abelian varieties over quadratic extensions}, invited speaker at {\it Rational points in Bristol}, University of Bristol, February 2018  \medskip \\
{\it Parity of Selmer ranks in quadratic twist families}, contributed talk for the conference {\it Curves and L-functions}, ICTP Trieste, September 2017  \medskip \\
{\it Parity of Selmer ranks in quadratic twist families}, Number Theory seminar, University of Manchester, February 2017  \medskip \\
{\it Parity of Selmer ranks in quadratic twist families}, Algebra seminar, University of Pennslyvania, November 2016  \medskip \\
{\it Parity of ranks of abelian varieties}, Quebec--Vermont Number Theory seminar, October 2016  \medskip \\
{\it Parity of Selmer ranks in quadratic twist families}, Number Theory seminar, University of Cambridge, September 2016  \medskip \\
{\it Parity of Selmer ranks in quadratic twist families}, contributed talk for the conference  {\it Arithmetic statistics and the Cohen--Lenstra heuristics}, University of Warwick, June 2016  \medskip \\
{\it 2-Selmer parity for Jacobians of hyperelliptic curves over quadratic extensions}, Geometry and Algebra seminar, Utrecht university, December 2015  \medskip \\
{\it 2-Selmer parity for Jacobians of hyperelliptic curves over quadratic extensions}, TCC Number Theory day, Imperial College London, April 2015  \medskip \\
%{\it Iwasawa theory of elliptic curves}, study group, University of Bristol, February 2015 \medskip \\
%{\it Bhargava's cubes; a modern perspective on Gauss composition}, Pure postgraduate seminar, University of Bristol, November 2014   \medskip \\
%{\it Intersection theory for surfaces}, study group, University of Bristol, August 2014   \medskip \\
%{\it The Weil conjectures for elliptic curves}, Study group, University of Bristol, August 2014   \medskip \\   
 {\it Parity of $2$-Selmer ranks of hyperelliptic curves over quadratic extensions}, contributed talk, CNTA XIII, Ottawa, June 2014  \medskip \\
 {\it On Bhargava and Shankar's work on the average rank of elliptic curves}, Linfoot seminar, University of Bristol, February 2014  \medskip \\
 {\it Parity of $2$-Selmer ranks of hyperelliptic curves over quadratic extensions}, University of Warwick Number Theory seminar, January 2014  \medskip \\
%{\it Sheaves of modules}, two talks for the TCC course `Algebraic Geometry for Number Theory', October-December 2013  \medskip \\
%{\it Modular forms and Galois representations}, four talks for a study group of the same name, University of Bristol, Oct-Dec 2013 \medskip  \\
%{\it Local class field theory}, Pure postgraduate seminar, University of Bristol, May 2013  \medskip \\
%{\it Hausdorff measure and dimension}, study group,  University of Bristol, October 2012 
       
%\section{EXPOSITORY WRITING}
%
%\bigskip
%
%  \vspace{-0.1in}	
%{\it Brauer groups and Galois cohomology}, essay for the TCC course `Arithmetic Geometry', January 2014, available on request.   \medskip \\
%{\it The Szemer\'{e}di-Trotter theorem}, essay for the TCC course `Additive Combinatorics', December 2013, available on request.

\section{TEACHING AND PROJECT SUPERVISION}       

\bigskip 
10 hour SMSTC lecture course {\it Algebras and Representation Theory}, University of Glasgow, Jan-Mar 2023.\medskip\\
12 hour SMSTC lecture course {\it Galois cohomology and central simple algebras}, University of Glasgow, Jan-Mar 2019.\medskip\\
Supervisor for two MSci projects {\it Parametrization of rings of small rank} 
\textup{and} {\it Composition of binary quadratic forms}\textup{, King's College London, June-Sept 2017} \medskip\\   
    {\it Linear algebra and geometry}, teaching assistant, University of Bristol, 2013-2014 \medskip\\         
    {\it Analysis}, teaching assistant, University of Bristol, 2013-2014 \medskip \\
{\it Group theory maths cafe}, organiser, University of Bristol, May 2013 \medskip \\
    {\it Maths 1A20} (introduction to calculus and analysis), teaching assistant, University of Bristol, 2012-2013. 
 

 
%\section{SELECTED CONFERENCES ATTENDED}          
%    \bigskip 
%    
%    {\it Workshop on Arithmetic Geometry, Number Theory, and Computation}, ICERM, June 2020 \medskip \\  
%     {\it  CMI-HIMR Summer School in Computational Number Theory}, University of Bristol, June 2019  \medskip \\
%    {\it Enumerative Arithmetic and the Cohen–Lenstra Heuristics}, MPIM, Bonn, June 2019 \medskip \\
%        {\it Arithmetic Geometry, Number Theory, and Computation}, MIT, August 2018 \medskip \\
%            {\it Arithmetic of Curves}, Workshop in Baskerville Hall, Wales, August  2018 \medskip \\
%            {\it Mathematics is a long conversation: 
%a celebration of Barry Mazur}, Harvard University, June 2018 \medskip \\
%                {\it Rational and Integral Points via Analytic and Geometric Methods}, CMO Oaxaca, June 2018 \medskip \\
%    {\it Curves and L-functions}, ICTP Trieste, September 2017 \medskip \\
%    {\it Perfectoid Spaces}, Arizona winter school, March 2017 \medskip \\
%{\it 2016 Fields medal symposium}, Fields Institute, Toronto, October 2016 \medskip \\
%{\it Arithmetic statistics and the Cohen--Lenstra heuristics}, University of Warwick, June 2016 \medskip \\
%{\it Explicit methods in number theory}, University of Warwick, April 2016 \medskip \\
%{\it LMS-CMI Summer school on Diophantine Equations}, Baskerville Hall, Hay-on-Wye, September 2015\medskip \\
%{\it Elliptic curves, modular forms and Iwasawa theory}, University of Cambridge, March 2015 \medskip \\
%{\it Arithmetic and higher dimensional varieties}, Arizona winter school, March 2015 \medskip \\
%{\it Counting arithmetic objects}, Summer school, Universit\'{e} de Montr\'{e}al, June-July 2014 \medskip \\
%{\it CNTA XIII}, Ottawa, June 2014 \medskip \\
%{\it Arithmetic Statistics}, Arizona winter school, March 2014 \medskip \\
%{\it Young researchers in mathematics}, University of Edinburgh, June 2013 \medskip \\
%{\it Rational Points - Geometric, Analytic and Explicit Approaches}, University of Warwick, May 2013 \medskip \\
%{\it Iwasawa Theory and Galois Representations}, University of Warwick, April 2013 \medskip \\
%{\it Explicit Methods for Modular Forms}, University of Warwick, March 2013 \medskip \\
%   {\it Selmer Groups, Descent and the Distribution of Ranks}, University of Warwick, September 2012 \medskip \\ 
%       
 
\end{resume}
\end{document}